---
title: "V and R"
subtitle: "rvee: recreational v programming for R" 
author: "Edwin de Jonge"
date: "Statistics Netherlands / UseR! 2022, @edwindjonge"
output:
  beamer_presentation:
    includes:
      in_header: header.tex
    keep_tex: yes
    theme: "Frankfurt"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(rvee)
```

##  {.plain .fragile}

```{=tex}
\begin{centering}
  \includegraphics[width=0.8\paperwidth]{img/rv}
  \par
\end{centering}
```
## Who am I?

### Edwin de Jonge

-   Statistical Consultant / R&D at Statistics Netherlands / CBS
-   Author of some R packages (e.g. `whisker`, `docopt`): <https://edwindj.r-universe.dev>
-   Love R and like programming in general
-   <http://github.com/edwindj>, @edwindjonge

## Short summary

\large

-   `V` new simple programming language

-   `rvee` : <https://github.com/edwindj/rvee> helps to create R extension in V.

-   No runtime and R dependencies.

-   Experimental, but works.

## What is V?

-   V is a programming language.

> Simple.fase, safe, compiled. 
For developing maintainable software.

-   Website: <https://vlang.io/>
-   young, low-level,
-   feels like scripting, but is strongly typed
-   Very similar to `Go`, but "improves" on Go: <https://vlang.io/compare#go>
-   Very light weight

## {.plain .fragile} Syntax example

```{=tex}
\begin{centering}
  \includegraphics[width=0.8\paperwidth]{img/example.v}
  \par
\end{centering}
```

## V is safe

```{=tex}
\includegraphics[width=0.1\paperwidth]{img/vlang}
```

- No `null`, `undefined` values: unexpected `null` causes software damage. 
- **immutable** variables by default, making it easier to optimize and reason about code
- **Option**: functions returns `?Type` and compiler generates error is empty return is not handled properly.
- **Result**: functions can return `!Type` and compiler generates error if run-time error is not handled properly.
- Other cases a function always returns a value
- Functions are pure by default (impurity is explicit)

## V features

-   Easy C-interop (C-backend)
-   Fast compilation!
-   Zero cost abstractions
-   Sum Types
-   Generics

## Fast compilation  {.plain .fragile} 
\begin{centering}
  \includegraphics[width=0.8\paperwidth]{img/compilation}
  \par
\end{centering}

\tiny source: <https://vlang.io>

## Performance (from V lang website)

-   As fast as C (V's main backend compiles to human readable C)
-   C interop without any costs
-   Minimal amount of allocations
-   Built-in serialization without runtime reflection
-   Compiles to native binaries without any dependencies: a simple web server is only 65 KB

### V has different compilation back-ends (like rust and Go)

-   native code (includes a REPL)
-   readable(!) `C` code
-   javascript

## R similarities

-   V functions are *pure* by default, like R, 
arguments are immutable by default. Have to be marked `mut`able.
- V has `defer`, like R's `on.exit`
- Simple syntax, very flexible
- Has (some) reflection/code generation. (R beats that one)

## V pkgs

-   V has a module and pkg system, containing `v` src code (not compiled)
-   Typical v-compilation results in one binary
-   importing a module means including the v code.
-   v can link though to `C`, `CPP` libraries.

## What is `rvee`?

- An experimental R package to create R package extensions with functions written in V.
- Similar in concept to RCpp but for V.
- But no _compile-time_ and _run-time_ dependencies in resulting R package.
- (pkgname `rv` was already claimed for)

### Why-o-why

- One of R's great powers is that of a glue language (just like S)
- V is a nice low-level language, so why not?
- Because it is fun :-)

## Features `rvee`
Two possible routes for using `v` as extension for R:
- Use v compiler to link to `R.so` / `R.dll`, but requires v to be present on CRAN.
- Use v transpiler to C and create a "normal" R package. This is `rvee`

### `rvee`

- Transpiles v code into C package code, including all wrappers
- include `vpkg`:`r` with wrapper functions for R's C API (30% complete, vector stuff works)

## `rvee::rv_export_c`

- Put the v source in `"<pkg>/src/v"` directory
- decorate each function to be exported with `[rv_export]` attribute
- `rvee::rv_export_c("<pkg>")` generates:

     - "./R/rv_export.R": R functions calling the v functions declared in "./src/v/rv_export.v"
     - "./src/v/rv_export.v": v wrapper functions translating input and output to the original v functions.
     - "./src/init.c": registration code for the shared library
     - "./src/<pkg>.c": the c code generated by v from the source files in the "<pkg>/src/v" directory.

- And ready! No dependencies what so ever.
- After that `devtools::load_all` (or `R CMD SHLIB`) work. (on the generated C files)

## Other `v_function`

- `v_function` generates from v src code a working R function that calls that code

```{r, eval=FALSE}
v_function('fn add(x int, y int, z int) int{
  sum := x + y + z
  return sum
}')

add(1,2,3)
```

### vpkg `r`

- wrapper library for R API.
- helps to transform V strings, arrays into R vectors (and vice versa).

## Status `rvee`

-   V is a young language, but already quite stable.
-   Generated code is in C
-   rvee is experimental,
-   Help is welcome!
-   Experimental, expect (breaking) changes!
-   But..., hey it works :-) 

## Thank you!

\Large Interested?

``` r
remotes::install_github("edwindj/rvee")
```

Or visit:

[**http://github.com/edwindj/rvee**](http://github.com/edwindj/rvee)
