---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `rvee`

## Recreational V programming for R / WIP

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/rvee)](https://CRAN.R-project.org/package=rvee)
<!-- badges: end -->

**Early work, expect errors and crashes! (not ready for production)**
 
Create R extension packages with the [V programming language](https://vlang.io).
V is a simple, safe and fast programming language with the speed of C.

R has many good interfaces with programming languages C, fortran, cpp (Rcpp), python
(reticulate), rust (), `rvee` provides this for the `v` programming language.

## Status

Translation to `C` and compilation:

Interfacing:

- [x] Generating all interfacing code from `v` file with `rvee::rv_export_c()`
- [x] wraps simple input and return types: `f64`, `int`, `string`
- [x] wraps `numeric`, `integer` and `character` input and return types.
- [x] compiling and working :-)
- [ ] CRAN checks (remove c compiler warnings)

r module (in v):

- [x] `Numeric`, direct access as a `[]f64`
- [x] `Integer`, direct access as a `[]int`
- [x] `Character`, indirect access as a `[]string`. string values of R are reused, but newly created string (not managed by R) are copied.
- [x] `Logical`, indirect access as a `[]bool`. automatically converts between `[]bool` and `logical`.
- [ ] `Factor` 
- [ ] `List`
- [ ] `DataFrame`
- [ ] `Environment`
- [ ] `Function`

## Installation


The **unstable** development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("edwindj/rvee")
```

## Overview

There are the following possible routes for creating an R extension with `v` code:

a) Transpiling `V` code to `C` code and use it as a normal C extension.
b) Using the `v` compiler to build a shared library linked to the R shared library

Both options have their benefits and draw backs:

a) Allows for easy distribution and installation, but requires tweaking the compilation
flags to pass CRAN checks etc. and removing/stripping code that is not needed by R.

b) Allows for a optimized shared library, but required `v` to be installed by the
installer (e.g. CRAN).

## Status

Currently the transpiling to C is in the make and kind-of works.

Put your `v` files in the "<pkg>/src" directory and decorate each function to be
exported with the `[rv_export]` attribute (see example below).

After that 

`rvee::rv_export_c` generates the necessary interfacing code:

- "./R/rv_export.R": R functions calling the v functions declared in "./src/rv_export.v"
- "./src/rv_export.v": v wrapper functions translating input and output to the original v functions.
- "./src/init.c": registration code for the shared library

And `devtools::load_all` (or `R CMD SHLIB`) should work.

Most work to be done is on the `v` module providing a easy interop between v and
R.

- `numeric` and `integer` vectors are working
- working on `character` vectors


## Example

Suppose we have the following file "<pkg>/src/example.v"

```{r, eval = FALSE, results="asis", code=readLines("example/example.v")}
```

With:

```{r, eval=FALSE}
rvee::rv_export_c("<pkg>", prefix="my_pkg") # <pkg> is root dir of the source of your package...
```
 
The interfacing code is generated:

`"<pkg>/src/rv_export.v"`:

```v
```{r, echo=FALSE, results='asis'}
fns <- rvee:::scan_v_file("example/example.v")
pkg <- "my_pkg"
rvee:::generate_rv_export_v(fns,pkg=pkg)
```
```

and 

`"<pkg>/R/rv_export.R"`:

```r
```{r, echo=FALSE, results='asis'}
rvee:::generate_rv_export_R(fns, pkg = pkg)
```
```
 
And create the shared library with a call to devtools

```{r, eval = FALSE}
devtools::load_all("<pkg>")
```
