---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `rvee`

## Recreational V programming for R / WIP

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/rvee)](https://CRAN.R-project.org/package=rvee)
<!-- badges: end -->

**Work in progress, not ready for production**

Create R extension packages with the [V programming language](https://vlang.io).
V is a simple, safe and fast programming language with the speed of C.

R has many good interfaces with programming languages C, fortran, cpp (Rcpp), python
(reticulate), rust ().

## Installation


The **unstable** development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("edwindj/rvee")
```

## Overview

There are the following possible routes for creating an R extension with `v` code:

a) Transpiling `V` code to `C` code and use it as a normal C extension.
b) Using the `v` compiler to build a shared library linked to the R shared library

Both options have their benefits and draw backs:

a) Allows for easy distribution and installation, but requires tweaking the compilation
flags to pass CRAN checks etc. and removing/stripping code that is not needed by R.

b) Allows for a optimized shared library, but required `v` to be installed by the
installer (e.g. CRAN).

## Status

Currently the transpiling to C is in the make and kind-of works.

Put your `v` files in the "<pkg>/src" directory and decorate each function to be
exported with the `[rv_export]` attribute (see example below).

After that 

`rvee::rv_export_c` generates the necessary interfacing code:

- "./R/rv_export.R": R functions interfacing the shared library
- "./src/rv_export.v": v wrapper functions translating input and output
- "./src/init.c": registration code for the shared library

And `devtools::load_all` (or `R CMD SHLIB`) should work.

Most work to be done is on the `v` module providing a easy interop between v and
R.

- `numeric` and `integer` vectors are working
- working on `character` vectors


## Example

Suppose we have the following file "<pkg>/src/example.v"

```{r, eval = FALSE, results="asis", code=readLines("example/example.v")}
```

With:

```{r, eval=FALSE}
rvee::rv_export_c("<pkg>") # <pkg> is root dir of the source of your package...
```

The interfacing code is generated:

`"<pkg>/src/rv_export.v"`:

```v
```{r, echo=FALSE, results='asis'}
fns <- rvee:::scan_v_file("example/example.v")
rvee:::generate_rv_export_v(fns)
```
```

and 

`"<pkg>/R/rv_export.R"`:

```r
```{r, echo=FALSE, results='asis'}
rvee:::generate_rv_export_R(fns)
```
```
