---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `rvee`

## Recreational V programming for R / WIP

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/rvee)](https://CRAN.R-project.org/package=rvee)
[![R-CMD-check](https://github.com/edwindj/rvee/workflows/R-CMD-check/badge.svg)](https://github.com/edwindj/rvee/actions)
<!-- badges: end -->

**Early work! (not ready for production)**
 
Create R extension packages with the [V programming language](https://vlang.io).
V is a simple, safe and fast programming language with the speed of C.

R has good interfaces for many programming languages such as [C](https://cran.r-project.org/doc/manuals/r-release/R-exts.html#System-and-foreign-language-interfaces)
, [fortran](https://cran.r-project.org/doc/manuals/r-release/R-exts.html#System-and-foreign-language-interfaces)
, cpp (e.g. [Rcpp](http://www.rcpp.org/))
, python (e.g. [reticulate](https://rstudio.github.io/reticulate/))
and rust (e.g. [r-rust](https://github.com/r-rust)).
R Package `rvee` intends to provide an easy toolkist for creating R packages with 
the `v` programming language.

## Status

Translation to `C` and compilation 🎉!

Interfacing:

- [x] Generating all interfacing code from `v` file with `rvee::rv_export_c()`
- [x] wraps simple input and return types: `f64`, `int`, `bool`,  `string`
- [x] wraps `numeric`, `logical`, `integer`, `character`, `factor` and `list` input and return types.
- [x] compiling and working :-)
- [ ] wraps `data.frame` input and return types.
- [ ] wraps array input and return types: `[]f64`, `[]int`, `[]bool`,  `[]string`
- [x] CRAN checks

r module (in v):

- [x] `Numeric`, direct access as a `[]f64`
- [x] `Integer`, direct access as a `[]int`
- [x] `Character`, indirect access as a `[]string`. string values of R are reused, but newly created string (not managed by R) are copied.
- [x] `Logical`, indirect access as a `[]bool`. automatically converts between `[]bool` and `logical`.
- [x] `Factor` 
- [x] `List`
- [x] `DataFrame`
- [ ] `Environment`
- [ ] `Function`

## Installation


The **unstable** development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("edwindj/rvee")
```

## Overview

Possible routes for creating an R extension with `v` code are:

a) Transpiling `V` code to `C` code and use it as a normal `C` extension.
b) Using the `v` compiler to build a shared library linked to the R shared library.

Both options have their benefits and draw backs:

a) Allows for easy distribution and installation, but requires tweaking the compilation
flags to pass CRAN checks etc. and removing/stripping code that is not needed by R.

b) Allows for an optimized shared library, but requires `v` to be installed by the
installer (e.g. CRAN).

## Example

Transpiling to C works.

Put your `v` files in the `"<pkg>/src/v"` directory and decorate each function to be
exported with the `[rv_export]` attribute (see example below).

After that 

`rvee::rv_export_c` generates the necessary interfacing code:

- `"./R/rv_export.R"`: R functions calling the v functions declared in `"./src/v/rv_export.v"`
- `"./src/v/rv_export.v"`: v wrapper functions translating input and output to the original v functions.
- `"./src/init.c"`: registration code for the shared library
- `"./src/<pkg>.c"`: the c code generated by `v` from the source files in the
`"<pkg>/src/v"` directory.

After that  `devtools::load_all` (or `R CMD SHLIB`) work: they will compile `"init.c"`
and `"<pkg>.c"` into `"<pkg>.so"`/`"<pkg>.dll"`.

Suppose we have the following file "<pkg>/src/example.v"

```{r, eval = FALSE, results="asis", code=readLines("example/example.v")}
```

With:

```{r, eval=FALSE}
rvee::rv_export_c("<pkg>", prefix="my_pkg") # <pkg> is root dir of the source of your package...
```
 
The interfacing code is generated:

`"<pkg>/src/v/rv_export.v"`:

```v
```{r, echo=FALSE, results='asis'}
fns <- rvee:::scan_v_file("example/example.v")
pkg <- "my_pkg"
rvee:::generate_rv_export_v(fns,pkg=pkg)
```
```

and 

`"<pkg>/R/rv_export.R"`:

```r
```{r, echo=FALSE, results='asis'}
rvee:::generate_rv_export_R(fns, pkg = pkg)
```
```
 
And create the shared library with a call to devtools

```{r, eval = FALSE}
devtools::load_all("<pkg>")
```
